---
title: "Bayesian Hierarchical Models for Personal Landspeeder Insurance"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggthemes)
library(brms)

train <- readRDS('./data/simulated_landspeeder_data_train.rds')
test <- readRDS('./data/simulated_landspeeder_data_test.rds')
```

### Let's fit some models!

First, a single GLM for everyone:

```{r}
mod_01_glm_all <-
    glm(total_claims ~ planet + pilot + vehicle,
        family = poisson(link = 'log'),
        data = train)

summary(mod_01_glm_all)
```

Now that we've built this model, let's look at the actual to expected to see
how well we predicted the outcome on our test data:

```{r}
# actual to expected by score decile
test %>%
    mutate(galaxywide_score = predict(mod_01_glm_all, test, type = 'response'),
           gw_score_decile = ntile(galaxywide_score, 20)) %>%
    group_by(gw_score_decile) %>%
    summarise(mean_claims = mean(total_claims),
              sigma_claims = sd(total_claims),
              mean_score = mean(galaxywide_score)) %>%
    ggplot(aes(x = gw_score_decile, y = mean_claims)) + 
    geom_line(color = 'blue', size = 1.5) + geom_point(color = 'blue') +
    geom_line(aes(y = mean_score), color = 'red') + 
    theme_tufte(base_size = 14,
                base_family = 'sans') + 
    theme(legend.title=element_blank()) +
    ggtitle('Galaxywide score, all planets')
```


Doesn't look too bad!  Let's see how well it does on each of our two planets:

```{r}
test %>% 
    mutate(galaxywide_score = predict(mod_01_glm_all, test, type = 'response'),
           gw_score_decile = ntile(galaxywide_score, 20)) %>%
    filter(planet == 'a.) Tatooine') %>%
    group_by(gw_score_decile) %>%
    summarise(mean_claims = mean(total_claims),
              sigma_claims = sd(total_claims),
              mean_score = mean(galaxywide_score)) %>%
    ggplot(aes(x = gw_score_decile, y = mean_claims)) + 
    geom_line(color = 'blue', size = 1.5) + geom_point(color = 'blue') +
    geom_line(aes(y = mean_score), color = 'red') + 
    theme_tufte(base_size = 14,
                base_family = 'sans') + 
    theme(legend.title=element_blank()) +
    ggtitle('Galaxywide score, Tatooine')
```

```{r}
test %>% 
    mutate(galaxywide_score = predict(mod_01_glm_all, test, type = 'response'),
           gw_score_decile = ntile(galaxywide_score, 20)) %>%
    filter(planet == 'b.) Hoth') %>%
    group_by(gw_score_decile) %>%
    summarise(mean_claims = mean(total_claims),
              sigma_claims = sd(total_claims),
              mean_score = mean(galaxywide_score)) %>%
    ggplot(aes(x = gw_score_decile, y = mean_claims)) + 
    geom_line(color = 'blue', size = 1.5) + geom_point(color = 'blue') +
    geom_line(aes(y = mean_score), color = 'red') + 
    theme_tufte(base_size = 14,
                base_family = 'sans') + 
    theme(legend.title=element_blank()) +
    ggtitle('Galaxywide score, Hoth')
```

Interesting.  What would happen if we fit a model to each planet independently?

```{r}
mod_02_glm_tatooine <-
    glm(total_claims ~ pilot + vehicle,
        family = poisson(link = 'log'),
        data = train[train$planet == 'a.) Tatooine', ])

summary(mod_02_glm_tatooine)
```

```{r}
mod_03_glm_hoth <-
    glm(total_claims ~ pilot + vehicle,
        family = poisson(link = 'log'),
        data = train[train$planet == 'b.) Hoth', ])

summary(mod_03_glm_hoth)
```

Hmm, it looks like the effect for the T-47 is completely different on Hoth as
it is on Tatooine.  The effects for Inexperienced pilots seems similar, but the 
p-values for these effects do not seem promising.  Let's take a closer look,
by looking at the actual mean claims vs the expected, group by group.

```{r}
train %>%
    transmute(vehicle, planet, total_claims,
              galaxywide_score = predict(mod_01_glm_all, type = 'response')) %>%
    gather(type, claim_rate, total_claims:galaxywide_score, factor_key = TRUE) %>%
    group_by(vehicle, type, planet) %>%
    summarise(mean_claims = mean(claim_rate),
              sigma = sd(claim_rate)) %>%
    ggplot(aes(x = vehicle, y = mean_claims, colour = type)) + 
    geom_line(aes(group = type), size = 1.5) +
    geom_pointrange(aes(ymin = mean_claims - sigma, 
                        ymax = mean_claims + sigma)) +  
    facet_grid(cols = vars(planet)) +
    theme_tufte(base_size = 16,
                base_family = 'sans') + 
    theme(legend.title=element_blank())
```

```{r}
train %>%
    transmute(pilot, planet, total_claims,
              galaxywide_score = predict(mod_01_glm_all, type = 'response')) %>%
    gather(type, claim_rate, total_claims:galaxywide_score, factor_key = TRUE) %>%
    group_by(pilot, type, planet) %>%
    summarise(mean_claims = mean(claim_rate),
              sigma = sd(claim_rate)) %>%
    ggplot(aes(x = pilot, y = mean_claims, colour = type)) + 
    geom_line(aes(group = type), size = 1.5) +
    geom_pointrange(aes(ymin = mean_claims - sigma, 
                        ymax = mean_claims + sigma)) +  
    facet_grid(rows = vars(planet)) + 
    theme_tufte(base_size = 16,
                base_family = 'sans') + 
    theme(legend.title=element_blank())
```

Let's try a multi-level model allowing the fixed effects to vary by planet:

```{r}
# mod_04_multilevel_planet <-
#     brm(total_claims ~ (1 | planet) + pilot + vehicle + 
#             (0 + pilot + vehicle | planet),
#         data = train, brmsfamily('poisson', link = 'log'),
#         chains = 1, iter = 100)
# 
# saveRDS(mod_04_multilevel_planet, './data/mod_04_multilevel_planet.rds')
mod_04_multilevel_planet <- readRDS('./data/mod_04_multilevel_planet_bg.rds')
```

```{r}
test_score_multilevel <- 
    predict(mod_04_multilevel_planet, newdata = test)
```

```{r}
test %>%
    transmute(vehicle, planet, total_claims,
              multilevel_bayes = test_score_multilevel[, 1]) %>%
    gather(type, claim_rate, total_claims:multilevel_bayes, factor_key = TRUE) %>%
    group_by(vehicle, type, planet) %>%
    summarise(mean_claims = mean(claim_rate),
              sigma = sd(claim_rate)) %>%
    ggplot(aes(x = vehicle, y = mean_claims, colour = type)) + 
    geom_line(aes(group = type), size = 1.5) +
    geom_pointrange(aes(ymin = mean_claims - sigma, 
                        ymax = mean_claims + sigma)) +  
    facet_grid(cols = vars(planet)) + 
    theme_tufte(base_size = 16,
                base_family = 'sans') + 
    theme(legend.title=element_blank())
```

```{r}
test %>%
    transmute(pilot, planet, total_claims,
              multilevel_bayes = test_score_multilevel[, 1]) %>%
    gather(type, claim_rate, total_claims:multilevel_bayes, factor_key = TRUE) %>%
    group_by(pilot, type, planet) %>%
    summarise(mean_claims = mean(claim_rate),
              sigma = sd(claim_rate)) %>%
    ggplot(aes(x = pilot, y = mean_claims, colour = type)) + 
    geom_line(aes(group = type), size = 1.5) +
    geom_pointrange(aes(ymin = mean_claims - sigma, 
                        ymax = mean_claims + sigma)) +  
    facet_grid(rows = vars(planet)) + 
    theme_tufte(base_size = 16,
                base_family = 'sans') + 
    theme(legend.title=element_blank())
```

```{r}
test %>% 
    mutate(multilevel_score = test_score_multilevel[, 1],
           score_decile = ntile(multilevel_score, 10)) %>%
    filter(planet == 'b.) Hoth') %>%
    group_by(score_decile) %>%
    summarise(mean_claims = mean(total_claims),
              sigma_claims = sd(total_claims),
              mean_score = mean(multilevel_score)) %>%
    ggplot(aes(x = score_decile, y = mean_claims)) + 
    geom_line(color = 'blue', size = 1.5) + geom_point(color = 'blue') +
    geom_line(aes(y = mean_score), color = 'red') + 
    theme_tufte(base_size = 14,
                base_family = 'sans') + 
    theme(legend.title=element_blank()) +
    ggtitle('Bayesian multilevel score, Hoth')
```

```{r}
test %>% 
    mutate(multilevel_score = test_score_multilevel[, 1],
           score_decile = ntile(multilevel_score, 20)) %>%
    group_by(score_decile) %>%
    summarise(mean_claims = mean(total_claims),
              sigma_claims = sd(total_claims),
              mean_score = mean(multilevel_score)) %>%
    ggplot(aes(x = score_decile, y = mean_claims)) + 
    geom_line(color = 'blue', size = 1.5) + geom_point(color = 'blue') +
    geom_line(aes(y = mean_score), color = 'red') + 
    theme_tufte(base_size = 14,
                base_family = 'sans') + 
    theme(legend.title=element_blank()) +
    ggtitle('Bayesian multilevel score, Tatooine')
```
